Space for WHAM: a multi-region, multi-stock generalization of the Woods Hole Assessment Model with an application to black sea bass

### 2025-01-21 paper/paper.Rmd
saving work

A state-space statistical approach and maximum marginal likelihood or Bayesian fitting of stock assessment models allows estimation of time- and age-varying population attributes as random effects [@nielsenberg14,@cadigan16,@milleretal16]. This estimation approach is considered an essential feature of gold-standard assessment models that we use in tactical management of commercially important fish stocks [@puntetal20]. The State-space Assessment Model [SAM, @nielsenberg14] continues to be developed and remains widely used within ICES to assess European fish stocks. Various state-space models are being used to manage cod and plaice stocks in the waters of Eastern Canada [@varkeyetal22;@perreaultetal20] and to the south, the Woods Hole Assessment Model [WHAM, @millerstock20; @stockmiller21] is now used to assess many fish stocks in the Northwest Atlantic Ocean [@nefsc22;@nefsc22a;@nefsc24].

>>>>

A state-space statistical approach and maximum marginal likelihood or Bayesian fitting of stock assessment models allows estimation of time- and age-varying population attributes as random effects [@nielsenberg14;@cadigan16;@milleretal16]. This estimation approach is considered an essential feature of gold-standard assessment models that we use in tactical management of commercially important fish stocks [@puntetal20]. The State-space Assessment Model [SAM, @nielsenberg14] continues to be developed and remains widely used within ICES to assess European fish stocks. Various state-space models are being used to manage cod and plaice stocks in the waters of Eastern Canada [@varkeyetal22;@perreaultetal20] and to the south, the Woods Hole Assessment Model [WHAM, @millerstock20; @stockmiller21] is now used to assess many fish stocks in the Northwest Atlantic Ocean [@nefsc22;@nefsc22a;@nefsc24].
